---
title: "Bioinformatics Class 16"
author: "Cesar Arcasi Matta"
date: "5/24/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TnSeq_practice 

Install package
```{r}
#install.packages("tidyverse")
```

Load libraries
```{r}
library(tidyr)

library(dplyr)
```

```{r}
library(ggplot2)

library(readr)
```

Load the file with experimental results from TnSeq experiment

```{r}
#Note: tbl_df function forwards the argument to "as_data_frame""

exp_data <- read.csv("Experimental_results.csv")

head(exp_data)
 
```

As you see, this table has several columns. Column “Strain” shows which strain was used in the experiment. “Environment” shows conditions in which microorganisms were propagated. Columns “Mut_ID” and “BC_ID” give information about type of mutations and barcodes that correspond to them. Columns “H_0”: “H_96” show frequenfy of each barcode over time.

Lets modify a table:

First, remove the column “X”. We are not going to use the information contained in it. For this we are going to use function “select"

#select() keeps only the variables you mention - Note: Select is an add on package of dplyr and not from R.

```{r}
# Remove column "X" from a dataframe

exp_data <- select(exp_data, - X)

head(exp_data)
```
 
 #Practice 
 
```{r}
Mut_str_env <- select(exp_data, Mut_ID, Strain, Environment)

head(Mut_str_env)
```
 
Lets get ready for plotting:

We are going to use ggplot2 package. Please check the syntax that is commonly used for plotting with this package. Does everything make sense?

# Information from a ggplot description page:
# ggplot(df, aes(x, y, <other aesthetics>))

To make a graph, we need two give a function 2 variables (2 columns) to plot them against each other.

What variables are we going to use?

Lets rearrange our table to be able to plot the data easily. Instead on keeping information about barcode frequency in rows, we are going to create a column “Time” with time points and a column “Frequency” with corresponding barcode frequencies.

ggplot(data = NULL, mapping = aes(), ..., en

ggplot() is used to construct the initial plot object, and is almost always followed by + to add component to the plot. There are three common ways to invoke ggplot:

ggplot(df, aes(x, y, <other aesthetics>))

ggplot(df)

ggplot()

WE HAVE TO MODIFY OUR TABLE SO THAT IN ONE COLUMN IS TIME AND THE OTHER IS FREQUENCY.

# First, check how function "gather" works

Gather takes multiple columns and collapses into key-value pairs, duplicating all other columns as needed. You use gather() when you notice that you have columns that are not variables.

gather(data, key = "key", value = "value", ..., na.rm = FALSE,
  convert = FALSE, factor_key = FALSE)
  
```{r}
exp_rearranged <-gather(exp_data, Generation, Frequency, H_0:H_96)

head(exp_rearranged)
```
  
You might have noticed that “Generation” column contains both “H” that stands for “hours” and numbers. Lets remove “H_” part from this column.

Check the syntax of “separate” function.

Given either regular expression or a vector of character positions, separate() turns a single character column into multiple columns.

separate(data, col, into, sep = "[^[:alnum:]]+", remove = TRUE,
  convert = FALSE, extra = "warn", fill = "warn", ...)

```{r}
# Separate values in "Generation" column into 2 columns

table_for_graph<-separate(exp_rearranged,Generation,into=c("H","Time"))

head(table_for_graph)

```

# Remove column "H" using function "select"

```{r}
table_for_graph<-select(table_for_graph, -H)

table_for_graph
```

You might have noticed that our table contains a lot of “NA” values. Go ahead and remove them with na.omit function. Don’t forget to check it’s syntax first!

?na.omit

These generic functions are useful for dealing with NAs in e.g., data frames. na.fail returns the object if it does not contain any missing values, and signals an error otherwise. na.omit returns the object with incomplete cases removed. na.pass returns the object unchanged.

na.omit(object, ...)

```{r}
table_cleaned <-na.omit(table_for_graph)

table_cleaned$Time <- as.numeric(table_cleaned$Time)

head(table_cleaned)

```

Now the table is ready. How are we going to plot all the values? Do we need to separate them in any way? If yes, then how?

```{r}
# We might need to plot data for each strain separately..

DivAnc<-filter(table_cleaned, table_cleaned$Strain=="DivAnc")

L013<-filter(table_cleaned, table_cleaned$Strain=="L013")
```

```{r}
# make a plot for DivAnc strain

DivAnc_plot=ggplot(DivAnc)+geom_line(aes(x=Time,y=Frequency,group=BC_ID),alpha=.2,colour="#000033")+ggtitle("DivAnc_SC3")+theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
DivAnc_plot

```

```{r}
# make a plot for L013 strain

L013_plot=ggplot(L013)+geom_line(aes(x=Time,y=Frequency,group=BC_ID),alpha=.2,colour="#CC6633")+ggtitle("L013_SC3")+theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
L013_plot
```

Can we make 2 graphs at the same time?

```{r}
ggplot(table_cleaned)+geom_line(aes(x=Time,y=Frequency,group=BC_ID),alpha=.2,colour="#000033")+facet_grid(.~Strain)+ggtitle("Barcode trajectories")+theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
```
Mutations from LO13 strain shows beneficial mutations that are positivly selected for in comparison to others.This is seen through a high mutation frequency, thus suggesting selection on mutation. DivAnc strain shows nuetral selection.

Lets pick one mutation and check how it behaves in different strains

filter = Use filter() find rows/cases where conditions are true. Unlike base subsetting with [, rows where the condition evaluates to NA are dropped.

== Binary operators which allow the comparison of values in atomic vectors.

```{r}
# I've chosen Mut_ID==34

mut34 <-filter(table_cleaned, table_cleaned$Mut_ID =="34")

mut34    
```

Plot Mut_ID == 34

```{r}
ggplot(mut34,aes(Time, Frequency, group=BC_ID, color=BC_ID))+geom_line()+theme(legend.position="none")+facet_grid(.~Strain)+ggtitle("Mutation_34")+xlab("Time, hours") + ylab("Log10(Barcode frequency)")+theme(plot.title = element_text(hjust = 0.5))
```

What can you tell looking at this plot? Why do we have 2 clusters of barcodes?
  -Two groups of barcodes, one on top with high frequency and the bottom with low. This may happen because the low frequeny coverage may represent poor barcode, thus unreliable due to its very low ferequncy. 

Lets filter out barcodes with frequency > (-5) and use them for subsequent analysis.

```{r}
mut34_f<-filter(mut34, mut34$Frequency>(-5))

mut34_f
```

Plot

```{r}
ggplot(mut34_f,aes(Time, Frequency, group=BC_ID, color=BC_ID))+geom_line()+theme(legend.position="none")+facet_grid(.~Strain)+ggtitle("Mutation_34")+xlab("Time, hours") + ylab("Log10(Barcode frequency)")+theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(mut34_f,aes(Time, Frequency, colour = BC_ID, group=BC_ID))+geom_point()+geom_smooth(se = FALSE, method = "lm")+facet_grid(.~Strain)+theme(legend.position="none")+ggtitle(paste("Mutation",34, sep="_"))+xlab("Time, hours")+ ylab("Log10(Barcode frequency)")
```

Now you chan choose a different mutation and check how it behaves across strains.

Now it’s time to estimate slope for each barcode. Lets greate a file that will contain information about BC_ID, Mut_ID, Strain, and estimated slope.

# Lets become familiar with lm function: lm is used to fit linear models. It can be used to carry out regression, single stratum analysis of variance and analysis of covariance (although aov may provide a more convenient interface for these).

# For this exercise, take the filtered data for mutation 34 (mut34_f) and filter out information about one barcode you like

# I have chosen BC_ID=25361 in DivAnc strain

```{r}
BC_25361 <- filter(mut34_f, mut34_f$BC_ID == "25361", mut34_f$Strain == "DivAnc")

BC_25361
```

#Lets plot frequency of this barcode:

```{r}
ggplot(BC_25361,aes(Time, Frequency, colour = BC_ID))+geom_point()+theme(legend.position="none")+ggtitle("BC_25361")+xlab("Time, hours") + ylab("Log10(Frequency)")
```

to be make it easy to ass trend line, you could name the ggplot code BC_plot and use BC_plot to add trend line. Ex. Code box #251

```{r}
BC_plot <- ggplot(BC_25361,aes(Time, Frequency, colour = BC_ID))+geom_point()+theme(legend.position="none")+ggtitle("BC_25361")+xlab("Time, hours") + ylab("Log10(Frequency)")
```

#Lets use lm function to fit the line to these points:

```{r}
BC_plot_lm <- BC_plot + geom_smooth(se = FALSE, method = "lm")

BC_plot_lm
```

# Lets check what data does lm function return:

Regression model = result of linear model function

```{r}
regression_model<-lm(Frequency~Time,BC_25361)

summary_data<-summary(regression_model)

summary_data
```

```{r}
# The information we are interested in, is the value of Slope and Intercept of this line:
# Let's try to access them:
# Time

Time<-summary_data$coefficients[2]

Time
```

```{r}
# Intercept:

Intercept<-summary_data$coefficients[1]

Intercept
```

Now we can find slopes for each barcode for each mutation in all strains.

```{r}
# Lets create the file:

data_header=matrix(data = NA,nrow = 1,ncol = 7)

data_header[1]="Mut_ID"

data_header[2]="BC_ID"        

data_header[3]="Strain"        

data_header[4]="Slope"

data_header[5]="Intercept"

data_header[6]="R^2"
```








